Propositions as Sessions, Mechanically

### 2015-04-13 paper/formalisation.tex
Tweakening the GV section.

The typing judgements for the functional language GV are shown in
Figure~\ref{fig:gv}. Note that I have altered the $n$-ary branch and choice
constructs to the restricted binary versions in order to avoid tricky
formulations. The change is not restrictive however, since CP itself provides
only binary versions of plus ($\oplus$) and with ($\with$)
constructs. Explicit limiting and unlimiting of abstraction was chosen to
enable the inclusion of these rules in the translation from GV to CP, which is
performed by induction on GV terms and is therefore
syntax-directed. Figure~\ref{fig:letscompare} presents the informal version of
the well-typed let rule. While \citeauthor{Char:2012} suggests a two-indices
approach for handling for multiple binders~\cite{Char:2012}, such a strategy
seems excessive for the GV let rule and the only rule to benefit from multiple
binders. For that reason, the simple solution of performing two openings was
chosen.

All types are amalgamated into one inductive definition and provide a
predicate for restricting typing rules to consider only valid session
types. As shown in \S~\ref{sec:fusion}, \citeauthor{Wadler:2014} defines GV
types as a set of mutually recursive definitions. Unfortunately, handling
mutually inductive definitions in Coq can be quite involved, requiring one to
either rely on the Coq system to provide a stronger mutual induction principle
or define one manually. Such a definition is possible, but it complicates
applications of the induction principle.

@@

Another issue related to types is how to represent type kinds, linear and
unlimited kind. Initially the type definition was indexed by the kind and
while this definition has the advantage of preventing certain ill-formed
types, such as a non-linear session type, it complicated later definitions and
proofs. For instance, the decidability of equality on types now relied on an
axiom permitting heterogeneous equality~\cite{McBride:1999}. Generally, it is
better to avoid unnecessary postulates if possible. Further, when formulating
the typing rules, the kind must be given to each type hypothesised and most
rules are polymorphic in the kind, requring an awkward kind inspection
function to be applied to each type within a binding. The alternative approach
taken was to define a well-formed type relation and remove the kind index from
the type definition. Figure~\ref{fig:wft} shows the well-formedness for the
linear function type. Notice, that even though the conclusion must have linear
kind, as in rule \coqe!wt_tm_labs!, the argument and result types can have any
kind, highlighting the aforementioned kind polymorphism. These issues
highlighted that one should be careful and considerate regarding when to use
dependent types. These issues highlighted that one should take care over the
representation and its effect on later expressiveness.

>>>>

The typing judgements for the functional language GV are shown in
Figure~\ref{fig:gv}. Note that I have altered the $n$-ary branch and choice
constructs to the restricted binary versions in order to avoid tricky
formulations. The change is not restrictive however, since CP itself provides
only binary versions of plus ($\oplus$) and with ($\with$)
constructs. Explicit limiting and unlimiting of abstraction was chosen to
enable the inclusion of these rules in the translation from GV to CP, which is
performed by induction on GV terms and is therefore
syntax-directed. Figure~\ref{fig:letscompare} presents the informal version of
the well-typed let rule. While \citeauthor{Char:2012} suggests using
two-indices for handling multiple binders~\cite{Char:2012}, such a strategy
seems excessive for the GV let rule and the only rule to benefit from multiple
binders. For that reason, the simple solution of performing two openings was
chosen.

As shown in \S~\ref{sec:fusion}, \citeauthor{Wadler:2014} defines GV types as
a set of mutually recursive definitions. Unfortunately, handling mutually
inductive definitions in Coq can be quite involved, requiring one to either
rely on the Coq system to provide a stronger mutual induction principle or
define one manually. Such a definition is possible, but it complicates
applications of the induction principle. Therefore, all types are amalgamated
into one inductive definition.

@@

Another issue related to types is how to represent type kinds, linear and
unlimited kind. Initially the type definition was indexed by the kind and
while this definition has the advantage of preventing certain ill-formed
types, such as a non-linear session type, it complicated later definitions and
proofs. For instance, the decidability of equality on types now relied on an
axiom permitting heterogeneous equality~\cite{McBride:1999}. Generally, it is
better to avoid unnecessary postulates if possible. Further, when formulating
the typing rules, the kind must be given to each type hypothesised and most
rules are polymorphic in the kind, requring an awkward kind inspection
function to be applied to each type within a binding. The alternative approach
taken was to define a well-formed type relation and remove the kind index from
the type definition. Figure~\ref{fig:wft} shows the well-formedness for the
linear function type. Notice, that even though the conclusion must have linear
kind, as in rule \coqe!wt_tm_labs!, the argument and result types can have any
kind, highlighting the aforementioned kind polymorphism. An additional
predicate was defined to ensure typing rules consider only valid session
types. The predicate appears in the well-formedness constructors for session
types, ensuring the continuation is a session type and not, for example, a
linear function type. These issues highlight that one should be mindful of the
chosen representation and its effect on later expressiveness.
